---
title: "Mada517_script"
author: "Talon Jost"
date: "2025-07-04"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## packages

```{r cars}
rm(list = ls())
library(tidyverse)
library(tidyr)
library(dplyr)
library(ape)
library(ggtree)
library(ggthemes)
library(ggtreeExtra)
library(DT)
library(magrittr)
library(ggplot2)
```

## Madagascar Culture Metadata input
```{r}

## merging Madagascar database metadata (culture info, etc) with Vsearch 100% clusterIDs

## sorted by seq plate, toxonomy string then isolate ID in excel before creation

MadaDBdata = read_tsv("./MadaCultureDB_Metadata.txt")

## This deals with the duplicate entries for the isolates; its not perfect, it just selects the first entry from the sorted file 

MadaDBdata_filt = MadaDBdata %>%
  group_by(Isolate.ID) %>% slice_head() %>% # filters to 1 sequence per group
  ungroup() %>%
  select(Isolate.ID,Frog.ID,`proportion growth 1`,`anti-Bd Class (80%)`,Taxnonomy,DirTax_Phylum,DirTax_Class,DirTax_Order,DirTax_Family,DirTax_Genus,DirTax_Species) # selecting only a subset of columns for now for easier viewing

```

## Vsearch data

```{r}
#(vsearch) mcbletz@Mollys-MacBook-Pro-2 ~ % vsearch --cluster_fast /Users/mcbletz/Library/CloudStorage/OneDrive-ThePennsylvaniaStateUniversity/Bletz_Lab/Computational_Resources/Datasets/CultureData/MadaCultureDB_seqs_dupsfilt.fasta --uc /Users/mcbletz/Library/CloudStorage/OneDrive-ThePennsylvaniaStateUniversity/Bletz_Lab/Computational_Resources/Datasets/CultureData/MadaCultureDB_seqs_clustered_per100_050725.txt --id 1 --strand both

##*** MODIFIED HEADERS AND COLUMNS MANUALLY IN EXCEL

## Vsearch done with fasta from MadaDB file filtered for duplicates; this has more isolates than AmphiBac; This is the vsearch output (code above); Dup filtering completed with MadaDB_VsearchClusterMergeCode.R file.

MadaDB_VsearchClusters2 = read_tsv("./MadaCultureDB_seqs_clustered_per100_050725.txt") %>%
  select(Record,Cluster,Isolate.ID,Id.Match) # selecting relevant columns


```

## Merging vsearch clusters with culture metadata

Isolate ID is the common identifier and is used to merge the two files

```{r}

## merging clustering with Mada DB metadata that has had the dubs filtered
## NOTE Vsearch was not done with the duplicate filtered fasta...I think that's OK for now.but it does mean that there could be isolates with same ID that are "different" clusters because there are some that didn't ID to the same taxonomy.
MadaDB_VsearchClusters_wdata = left_join(MadaDB_VsearchClusters2,
                                         MadaDBdata_filt,
                                         by = "Isolate.ID") 
#write_tsv(MadaDB_VsearchClusters_wdata,"./MadaCultureDB_Metadata_wVsearchClusters.txt")
```

## Nesting

Nesting groups the Vsearch clusters for further summarizing

IMPORTANT - THERE ARE actually 1387 unique clusters when including singletons!!
There are 817 when we apply the "extreme" function filter

```{r}

MadaDB_VsearchClusters_wdata_nest = MadaDB_VsearchClusters_wdata %>%
  filter(!Record=="C") %>% # removing these records that are not important for the clustering its a duplicate of the S ids; only "S" and "H" retained
  filter(`anti-Bd Class (80%)` == "Inhibitory" | `anti-Bd Class (80%)` == "NonInhibitory") %>%
  filter(`proportion growth 1` > 0.75 | `proportion growth 1` < 0.25) %>% #added this to filter to extreme function
  group_by(Cluster) %>%
  nest()

```

## Summarizing

This creates an initial summary of the functional categories for each cluster


```{r}

summary_fxn = function(x){
  x %>% ## called x because thats what the position in the map function is called
  group_by(`anti-Bd Class (80%)`) %>%
  summarize(Count = n()) %>%
    pivot_wider(.,names_from = `anti-Bd Class (80%)`,values_from = Count)

} # this is a function that does the summarizing by function category and then organizes it for better viewing.


MadaDB_VsearchClusters_wdata_nest517 = MadaDB_VsearchClusters_wdata %>%
  filter(!Record=="C") %>% # removing these records that are not important for the clustering its a duplicate of the S ids; only "S" and "H" retained
  #filter(`anti-Bd Class (80%)` == "Inhibitory" | `anti-Bd Class (80%)` == "NonInhibitory") %>%
  group_by(Cluster) %>%
  filter(!Id.Match == "*") %>%  ## Why? This actually gets rid of singletons
  nest() 
  
  ## List of IDs for tree for all cluster dataset
MadaDB_VsearchClusters_wdata_nestnosingles_IDs = MadaDB_VsearchClusters_wdata_nest517 %>%
  unnest(data) %>%
  select(Id.Match) %>%
  filter(!Id.Match == "*") %>% ## This is needed to remove the creation of an "asterisk" group
  unique()

### STOPPED HERE DURING MEETING on 17 June 2025 ############## 



#MadaDB_VsearchClusters_wdata_nest2 = MadaDB_VsearchClusters_wdata_nest %>%
 # mutate(TotalIso = map(data,~ nrow(.x))) %>% # filter clusters with 1 isolate
  #filter(TotalIso > 2) %>% 
  #unnest(data) %>%
  #group_by(DirTax_Phylum) %>%
  #summarize(Count = n_distinct(Cluster))


MadaDB_VsearchClusters_wdata_nest517_fxn = MadaDB_VsearchClusters_wdata_nest517 %>%
  mutate(TotalIso = map(data,~ nrow(.x)))%>%
  mutate(FunctionCount = map(data,summary_fxn)) %>% # this creates the function categories
  unnest(FunctionCount) %>% # this allows the function categories to be viewed in the nested frame and "operated" on
  rename(Null = `NA`)
 

```




# 517 Molly version
```{r}

## needs files from above meta.data and tree.original and nest517

MadaDB_VsearchClusters_wdata_nest517 = MadaDB_VsearchClusters_wdata_nest517

meta.data <- MadaDBdata


tree.original <- ape::read.tree('Trees/tree/nwk/madaDB.rooted.tree.nwk')

MadaDB_517_IDs <- MadaDB_VsearchClusters_wdata_nestnosingles_IDs

## Molly addition 5/21/25 - using functions from above 
MadaDB_VsearchClusters_wdata_nest517_fxn = MadaDB_VsearchClusters_wdata_nest517 %>%
  mutate(TotalIso = map(data,~ nrow(.x)))%>%
  mutate(FunctionCount = map(data,summary_fxn)) %>% # this creates the function categories
  unnest(FunctionCount) %>% # this allows the function categories to be viewed in the nested frame and "operated" on
  rename(Null = `NA`)


mada517MB <- MadaDB_VsearchClusters_wdata_nest517_fxn %>% 
  dplyr::mutate(
    prop.inhib = as.numeric(Inhibitory) / as.numeric(TotalIso) * 100,
    prop.noninhib = as.numeric(NonInhibitory) / as.numeric(TotalIso) * 100,
    prop.na = as.numeric(Null) / as.numeric(TotalIso) * 100,
    prop.contam = as.numeric(Contam) / as.numeric(TotalIso) * 100) %>%
  left_join(MadaDB_517_IDs, by = "Cluster") %>% 
  dplyr::rename(Isolate.IDb = "Id.Match")



mada517.longMB <- mada517MB %>%
  select(Isolate.IDb,prop.inhib, prop.noninhib, prop.na, prop.contam) %>%
  pivot_longer(
    cols = c(prop.inhib, prop.noninhib, prop.na, prop.contam),
    names_to = "type",
    values_to = "percentage"
  )

#bar.long$Isolate.ID <- factor(bar.long$Isolate.ID, levels = rev(tree.pruned$tip.label))



full.tree2 <- ggtree(tree.original, 
                    branch.length = "none",
                    size = 0.5) %<+% meta.data +
  geom_tree(aes(color = DirTax_Class))

#full.tree2
#ggsave("full_tree.pdf", full.tree, height = 10, width = 15, dpi = 1000)


# updated
keep.517MB <- intersect(mada517MB$Isolate.IDb, tree.original$tip.label)

tree.pruned.517MB <- keep.tip(tree.original, keep.517MB)
plot(tree.pruned.517MB)

#tip_order <- tree.pruned.517MB$tip.label 

#tip_order_df <- tree.pruned.517MB$tip.label %>% as.data.frame() %>%
 # mutate(OrderNum = row_number())

#colnames(tip_order_df) <- c("Isolate.IDb", "OrderNum")

tree517MB <- ggtree(tree.pruned.517MB, 
       branch.length = "none", 
       size = 0.5) %<+% meta.data +
  geom_tree(aes(color = DirTax_Class)) +
  geom_tiplab(size = 5)

tree517MB

#ggsave("517_treeMB.pdf", tree517MB, height = 90, width = 30, dpi = 600, limitsize = FALSE)

## check point for number of clusters/branches
length(unique(mada517.longMB$Isolate.IDb))

## this are the lists to make the bar chart data be the same order as the tree;  Note tree order when the plot is printer is NOT what is extracted from tip label functions....
tree_order <- c("Mada.2840","Mada.3145","Mada.2715","Mada.1299","Mada.1116","Mada.2297","Mada.1808","Mada.3553","Mada.3378","Mada.3023","Mada.3407","Mada.4185","Mada.2377","Mada.2105","Mada.3559","Mada.2343","Mada.3625","Mada.3783","Mada.221","Mada.3759","Mada.1298","Mada.3615","Mada.2534","Mada.1844","Mada.3775","Mada.2846","Mada.3244","Mada.265","Mada.1518","Mada.4033","Mada.2370","Mada.2269","Mada.2340","Mada.3466","Mada.241","Mada.3902","Mada.1198","Mada.236","Mada.3294","Mada.3747","Mada.1837","Mada.4036","Mada.3539","Mada.2375","Mada.350","Mada.1564","Mada.3093","Mada.3579","Mada.3222","Mada.2350","Mada.3853","Mada.1263","Mada.834","Mada.4164","Mada.3289","Mada.2492","Mada.1712","Mada.918","Mada.2784","Mada.346","Mada.3346","Mada.1","Mada.2476","Mada.2525","Mada.2378","Mada.1296","Mada.3586","Mada.1816","Mada.3386","Mada.3265","Mada.3345","Mada.2439","Mada.2578","Mada.2853","Mada.2819","Mada.3735","Mada.3825","Mada.1926","Mada.2012","Mada.3419","Mada.2892","Mada.466","Mada.2307","Mada.3290","Mada.1752","Mada.4332","Mada.2969","Mada.3071","Mada.3821","Mada.3227","Mada.4190","Mada.1183","Mada.1301","Mada.3550","Mada.38","Mada.2895","Mada.2783","Mada.2951","Mada.2420","Mada.68","Mada.3777","Mada.3953","Mada.2181","Mada.2563","Mada.2353","Mada.2740","Mada.4132","Mada.1952","Mada.2364","Mada.3996","Mada.4180","Mada.3714","Mada.1342","Mada.216","Mada.2982","Mada.3597","Mada.2346","Mada.1084","Mada.2413","Mada.3091","Mada.1196","Mada.4334","Mada.2154","Mada.2816","Mada.2947","Mada.1127","Mada.349","Mada.3567","Mada.2824","Mada.2882","Mada.3676","Mada.3394","Mada.3465","Mada.2500","Mada.124","Mada.3712","Mada.4245","Mada.2402","Mada.4193","Mada.134","Mada.117","Mada.2579","Mada.3189","Mada.494","Mada.2768","Mada.229","Mada.434","Mada.1842","Mada.1119","Mada.4128","Mada.2851","Mada.2681","Mada.542","Mada.3321","Mada.1147","Mada.2667","Mada.4099","Mada.1359","Mada.4123","Mada.1954","Mada.3762","Mada.3024","Mada.180","Mada.740","Mada.3504","Mada.1793","Mada.2434","Mada.3224","Mada.451","Mada.1883","Mada.3501","Mada.3342","Mada.671")

tree_order2 <- c("Mada.412","Mada.3223","Mada.2436","Mada.1907","Mada.1249","Mada.1739","Mada.1273","Mada.2827","Mada.3343","Mada.298","Mada.648","Mada.1649","Mada.2543","Mada.1516","Mada.2431","Mada.3372","Mada.2427","Mada.1283","Mada.3081","Mada.168","Mada.3218","Mada.1418","Mada.2573","Mada.3200","Mada.2417","Mada.3532","Mada.3254","Mada.3866","Mada.2599","Mada.403","Mada.2393","Mada.3144","Mada.2894","Mada.2871","Mada.3601","Mada.1187","Mada.4158","Mada.3760","Mada.255","Mada.3418","Mada.4154","Mada.2702","Mada.2709","Mada.553","Mada.1766","Mada.3352","Mada.3362","Mada.2507","Mada.729","Mada.2488","Mada.3274","Mada.2443","Mada.3541","Mada.4024","Mada.4236","Mada.4234","Mada.1235","Mada.2424","Mada.3958","Mada.1700","Mada.2204","Mada.2914","Mada.3964","Mada.3017","Mada.2435","Mada.2349","Mada.4171","Mada.3359","Mada.2173","Mada.1759","Mada.2423","Mada.2897","Mada.2397","Mada.3279","Mada.2891","Mada.2757","Mada.3705","Mada.4065","Mada.2848","Mada.4012","Mada.3788","Mada.2758","Mada.3646","Mada.2411","Mada.3738","Mada.3694","Mada.2267","Mada.2383","Mada.2408","Mada.3175","Mada.4233","Mada.454","Mada.2365","Mada.3382","Mada.2385","Mada.2805","Mada.4320","Mada.97","Mada.2643","Mada.731","Mada.930","Mada.1826","Mada.2916","Mada.2682","Mada.1849","Mada.2577","Mada.3652","Mada.3528","Mada.929","Mada.2425","Mada.2741","Mada.2832","Mada.758","Mada.2679","Mada.387","Mada.739","Mada.2905","Mada.811","Mada.2723","Mada.2442","Mada.3622","Mada.1002","Mada.1783","Mada.2176","Mada.235","Mada.3153","Mada.1853","Mada.3555","Mada.3456","Mada.2287","Mada.2482","Mada.239","Mada.2618","Mada.1194","Mada.3438","Mada.1216","Mada.250","Mada.2904","Mada.351","Mada.3232","Mada.4242","Mada.3583","Mada.4179","Mada.4189","Mada.3213","Mada.3190","Mada.3830","Mada.3534","Mada.2215","Mada.2733","Mada.3843","Mada.1228","Mada.3458","Mada.2902","Mada.449","Mada.3184","Mada.3835","Mada.2316","Mada.1189","Mada.136","Mada.1553","Mada.1343","Mada.212","Mada.4328","Mada.211","Mada.3416","Mada.195","Mada.3828","Mada.2909","Mada.452","Mada.2549","Mada.397","Mada.1537","Mada.2678","Mada.3428","Mada.1792","Mada.3443","Mada.3847","Mada.3054","Mada.292","Mada.4131","Mada.2765","Mada.4209","Mada.3399","Mada.3854","Mada.4237","Mada.2024","Mada.984","Mada.3076","Mada.122","Mada.2441","Mada.1193","Mada.3406","Mada.4015","Mada.2468","Mada.2669","Mada.3479","Mada.3058","Mada.3852","Mada.1166","Mada.2586","Mada.3130","Mada.1873","Mada.3078","Mada.32","Mada.3075","Mada.2187","Mada.53","Mada.126","Mada.723","Mada.4009","Mada.2673","Mada.65","Mada.1888","Mada.2710","Mada.3920","Mada.4050","Mada.2373","Mada.34","Mada.2133","Mada.2004","Mada.457","Mada.157","Mada.2652","Mada.3159","Mada.3464","Mada.4035","Mada.3354","Mada.2962","Mada.2529","Mada.2119","Mada.2243","Mada.2523","Mada.1495","Mada.2600","Mada.3933","Mada.2400","Mada.155","Mada.160","Mada.290","Mada.1858","Mada.66","Mada.2314","Mada.637","Mada.1559","Mada.2654","Mada.2644","Mada.2950","Mada.2557","Mada.2102","Mada.3917","Mada.3166","Mada.3117","Mada.1358","Mada.3932","Mada.3404","Mada.716","Mada.1561","Mada.2324","Mada.283","Mada.3063","Mada.2023","Mada.3473","Mada.3127","Mada.2540","Mada.3309","Mada.166","Mada.1523","Mada.3272","Mada.3476","Mada.581","Mada.2226","Mada.118","Mada.1515","Mada.3537","Mada.3707","Mada.4306","Mada.41","Mada.3112","Mada.2842","Mada.2866","Mada.1007","Mada.384","Mada.3162","Mada.2548","Mada.100","Mada.3485","Mada.3666","Mada.1497","Mada.3726","Mada.4301","Mada.4313","Mada.407","Mada.2362","Mada.2339","Mada.3684","Mada.4146","Mada.1865","Mada.2619","Mada.1817","Mada.2608","Mada.2394","Mada.3413","Mada.184","Mada.3674","Mada.2604","Mada.3851","Mada.3796","Mada.1327","Mada.423","Mada.392","Mada.1365","Mada.2605","Mada.2046","Mada.1986","Mada.1251","Mada.1254","Mada.1670","Mada.2437","Mada.3496","Mada.2814","Mada.1905","Mada.614","Mada.2325","Mada.2641","Mada.2401","Mada.3695","Mada.4007","Mada.1852","Mada.3209","Mada.2850","Mada.1132","Mada.2301","Mada.2367","Mada.2756","Mada.715","Mada.1965","Mada.3383","Mada.2387","Mada.3034","Mada.1716","Mada.3326","Mada.2591","Mada.2637")

tree_order_all = c(tree_order, tree_order2)

# reversing order because the text string was top to bottom and tree/ bar plot is made bottom to top.
tree_order_all_rev = rev(tree_order_all)

barchart.full.mada517MB <- mada517.longMB %>%
  ggplot(aes(x = percentage, 
             y = factor(Isolate.IDb, levels = tree_order_all_rev), fill = type)) + ## ** this line is key
  #ggplot(aes(x = percentage, y = as.factor(OrderNum), fill = type)) +
  geom_bar(stat = "identity", position = "stack", width = 0.5) +
  ggthemes::scale_fill_colorblind()+
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_text(size = 10),
    legend.key.size = unit(2, "lines"),
    legend.title = element_text(size = 25),
    legend.text = element_text(size = 25)
  )


barchart.full.mada517MB
#ggsave("./517MB_barchart_factorlevels.pdf", barchart.full.mada517MB, height = 150, width = 30, dpi = 600, limitsize = FALSE)




```


#517 TBJ plus 77 version
```{r}

## needs files from above meta.data and tree.original and nest517

MadaDB_VsearchClusters_wdata_nest517 = MadaDB_VsearchClusters_wdata_nest517

meta.data = meta.data

tree.original = tree.original

MadaDB_517_IDs <- MadaDB_VsearchClusters_wdata_nestnosingles_IDs

## Molly addition 5/21/25 - using functions from above 
MadaDB_VsearchClusters_wdata_nest517_fxn = MadaDB_VsearchClusters_wdata_nest517 %>%
  mutate(TotalIso = map(data,~ nrow(.x)))%>%
  mutate(FunctionCount = map(data,summary_fxn)) %>% # this creates the function categories
  unnest(FunctionCount) %>% # this allows the function categories to be viewed in the nested frame and "operated" on
  rename(Null = `NA`)

MadaDB_VsearchClusters_wdata_nest517_IDs = MadaDB_VsearchClusters_wdata_nest517 %>%
  unnest(data) %>%
  select(Id.Match) %>%
  filter(!Id.Match == "*") %>% 
  unique()

mada517MB <- MadaDB_VsearchClusters_wdata_nest517_fxn %>% 
  dplyr::mutate(
    prop.inhib = as.numeric(Inhibitory) / as.numeric(TotalIso) * 100,
    prop.noninhib = as.numeric(NonInhibitory) / as.numeric(TotalIso) * 100,
    prop.na = as.numeric(Null) / as.numeric(TotalIso) * 100,
    prop.contam = as.numeric(Contam) / as.numeric(TotalIso) * 100) %>%
  left_join(MadaDB_VsearchClusters_wdata_nest517_IDs, by = "Cluster") %>% 
  dplyr::rename(Isolate.IDb = "Id.Match")



mada517.longMB <- mada517MB %>%
  select(Isolate.IDb,prop.inhib, prop.noninhib, prop.na, prop.contam) %>%
  pivot_longer(
    cols = c(prop.inhib, prop.noninhib, prop.na, prop.contam),
    names_to = "type",
    values_to = "percentage"
  )

#bar.long$Isolate.ID <- factor(bar.long$Isolate.ID, levels = rev(tree.pruned$tip.label))




#ggsave("full_tree.pdf", full.tree, height = 10, width = 15, dpi = 1000)


full.tree2.1 <- ggtree(tree.original, 
                     branch.length = "none",
                     size = 0.5) %<+% meta.data +
  geom_tree(aes(color = DirTax_Class)) +
  scale_color_manual(values = c(
    "Actinobacteria" = "#4E79A7",     
    "Deinococci" = "#F28E2B",    
    "Alphaproteobacteria" = "#E15759",         
    "Gammaproteobacteria" = "#76B7B2",
    "BetaproteoBacteri" = "#59A14F",
    "Flavobacteriia" = "#EDC948",
    "Sphingobacteriia" = "#B07AA1",
    "Bacilli" = "#FF9DA7",
    "NA" = "#BAB0AC"
  ))

full.tree2.1

# updated
keep.517MB <- intersect(mada517MB$Isolate.IDb, tree.original$tip.label)

tree.pruned.517MB <- keep.tip(tree.original, keep.517MB)
plot(tree.pruned.517MB)

#tip_order <- tree.pruned.517MB$tip.label 

#tip_order_df <- tree.pruned.517MB$tip.label %>% as.data.frame() %>%
 # mutate(OrderNum = row_number())

#colnames(tip_order_df) <- c("Isolate.IDb", "OrderNum")

tree517MB <- ggtree(tree.pruned.517MB, 
       branch.length = "none", 
       size = 0.5) %<+% meta.data +
  geom_tree(aes(color = DirTax_Class)) +
  geom_tiplab(size = 5)

tree517MB

#ggsave("517_treeMB.pdf", tree517MB, height = 90, width = 30, dpi = 600, limitsize = FALSE)

## check point for number of clusters/branches
length(unique(mada517.longMB$Iso))

## this are the lists to make the bar chart data be the same order as the tree;  Note tree order when the plot is printer is NOT what is extracted from tip label functions....
tree_order <- c("Mada.2840","Mada.3145","Mada.2715","Mada.1299","Mada.1116","Mada.2297","Mada.1808","Mada.3553","Mada.3378","Mada.3023","Mada.3407","Mada.4185","Mada.2377","Mada.2105","Mada.3559","Mada.2343","Mada.3625","Mada.3783","Mada.221","Mada.3759","Mada.1298","Mada.3615","Mada.2534","Mada.1844","Mada.3775","Mada.2846","Mada.3244","Mada.265","Mada.1518","Mada.4033","Mada.2370","Mada.2269","Mada.2340","Mada.3466","Mada.241","Mada.3902","Mada.1198","Mada.236","Mada.3294","Mada.3747","Mada.1837","Mada.4036","Mada.3539","Mada.2375","Mada.350","Mada.1564","Mada.3093","Mada.3579","Mada.3222","Mada.2350","Mada.3853","Mada.1263","Mada.834","Mada.4164","Mada.3289","Mada.2492","Mada.1712","Mada.918","Mada.2784","Mada.346","Mada.3346","Mada.1","Mada.2476","Mada.2525","Mada.2378","Mada.1296","Mada.3586","Mada.1816","Mada.3386","Mada.3265","Mada.3345","Mada.2439","Mada.2578","Mada.2853","Mada.2819","Mada.3735","Mada.3825","Mada.1926","Mada.2012","Mada.3419","Mada.2892","Mada.466","Mada.2307","Mada.3290","Mada.1752","Mada.4332","Mada.2969","Mada.3071","Mada.3821","Mada.3227","Mada.4190","Mada.1183","Mada.1301","Mada.3550","Mada.38","Mada.2895","Mada.2783","Mada.2951","Mada.2420","Mada.68","Mada.3777","Mada.3953","Mada.2181","Mada.2563","Mada.2353","Mada.2740","Mada.4132","Mada.1952","Mada.2364","Mada.3996","Mada.4180","Mada.3714","Mada.1342","Mada.216","Mada.2982","Mada.3597","Mada.2346","Mada.1084","Mada.2413","Mada.3091","Mada.1196","Mada.4334","Mada.2154","Mada.2816","Mada.2947","Mada.1127","Mada.349","Mada.3567","Mada.2824","Mada.2882","Mada.3676","Mada.3394","Mada.3465","Mada.2500","Mada.124","Mada.3712","Mada.4245","Mada.2402","Mada.4193","Mada.134","Mada.117","Mada.2579","Mada.3189","Mada.494","Mada.2768","Mada.229","Mada.434","Mada.1842","Mada.1119","Mada.4128","Mada.2851","Mada.2681","Mada.542","Mada.3321","Mada.1147","Mada.2667","Mada.4099","Mada.1359","Mada.4123","Mada.1954","Mada.3762","Mada.3024","Mada.180","Mada.740","Mada.3504","Mada.1793","Mada.2434","Mada.3224","Mada.451","Mada.1883","Mada.3501","Mada.3342","Mada.671")

tree_order2 <- c("Mada.412","Mada.3223","Mada.2436","Mada.1907","Mada.1249","Mada.1739","Mada.1273","Mada.2827","Mada.3343","Mada.298","Mada.648","Mada.1649","Mada.2543","Mada.1516","Mada.2431","Mada.3372","Mada.2427","Mada.1283","Mada.3081","Mada.168","Mada.3218","Mada.1418","Mada.2573","Mada.3200","Mada.2417","Mada.3532","Mada.3254","Mada.3866","Mada.2599","Mada.403","Mada.2393","Mada.3144","Mada.2894","Mada.2871","Mada.3601","Mada.1187","Mada.4158","Mada.3760","Mada.255","Mada.3418","Mada.4154","Mada.2702","Mada.2709","Mada.553","Mada.1766","Mada.3352","Mada.3362","Mada.2507","Mada.729","Mada.2488","Mada.3274","Mada.2443","Mada.3541","Mada.4024","Mada.4236","Mada.4234","Mada.1235","Mada.2424","Mada.3958","Mada.1700","Mada.2204","Mada.2914","Mada.3964","Mada.3017","Mada.2435","Mada.2349","Mada.4171","Mada.3359","Mada.2173","Mada.1759","Mada.2423","Mada.2897","Mada.2397","Mada.3279","Mada.2891","Mada.2757","Mada.3705","Mada.4065","Mada.2848","Mada.4012","Mada.3788","Mada.2758","Mada.3646","Mada.2411","Mada.3738","Mada.3694","Mada.2267","Mada.2383","Mada.2408","Mada.3175","Mada.4233","Mada.454","Mada.2365","Mada.3382","Mada.2385","Mada.2805","Mada.4320","Mada.97","Mada.2643","Mada.731","Mada.930","Mada.1826","Mada.2916","Mada.2682","Mada.1849","Mada.2577","Mada.3652","Mada.3528","Mada.929","Mada.2425","Mada.2741","Mada.2832","Mada.758","Mada.2679","Mada.387","Mada.739","Mada.2905","Mada.811","Mada.2723","Mada.2442","Mada.3622","Mada.1002","Mada.1783","Mada.2176","Mada.235","Mada.3153","Mada.1853","Mada.3555","Mada.3456","Mada.2287","Mada.2482","Mada.239","Mada.2618","Mada.1194","Mada.3438","Mada.1216","Mada.250","Mada.2904","Mada.351","Mada.3232","Mada.4242","Mada.3583","Mada.4179","Mada.4189","Mada.3213","Mada.3190","Mada.3830","Mada.3534","Mada.2215","Mada.2733","Mada.3843","Mada.1228","Mada.3458","Mada.2902","Mada.449","Mada.3184","Mada.3835","Mada.2316","Mada.1189","Mada.136","Mada.1553","Mada.1343","Mada.212","Mada.4328","Mada.211","Mada.3416","Mada.195","Mada.3828","Mada.2909","Mada.452","Mada.2549","Mada.397","Mada.1537","Mada.2678","Mada.3428","Mada.1792","Mada.3443","Mada.3847","Mada.3054","Mada.292","Mada.4131","Mada.2765","Mada.4209","Mada.3399","Mada.3854","Mada.4237","Mada.2024","Mada.984","Mada.3076","Mada.122","Mada.2441","Mada.1193","Mada.3406","Mada.4015","Mada.2468","Mada.2669","Mada.3479","Mada.3058","Mada.3852","Mada.1166","Mada.2586","Mada.3130","Mada.1873","Mada.3078","Mada.32","Mada.3075","Mada.2187","Mada.53","Mada.126","Mada.723","Mada.4009","Mada.2673","Mada.65","Mada.1888","Mada.2710","Mada.3920","Mada.4050","Mada.2373","Mada.34","Mada.2133","Mada.2004","Mada.457","Mada.157","Mada.2652","Mada.3159","Mada.3464","Mada.4035","Mada.3354","Mada.2962","Mada.2529","Mada.2119","Mada.2243","Mada.2523","Mada.1495","Mada.2600","Mada.3933","Mada.2400","Mada.155","Mada.160","Mada.290","Mada.1858","Mada.66","Mada.2314","Mada.637","Mada.1559","Mada.2654","Mada.2644","Mada.2950","Mada.2557","Mada.2102","Mada.3917","Mada.3166","Mada.3117","Mada.1358","Mada.3932","Mada.3404","Mada.716","Mada.1561","Mada.2324","Mada.283","Mada.3063","Mada.2023","Mada.3473","Mada.3127","Mada.2540","Mada.3309","Mada.166","Mada.1523","Mada.3272","Mada.3476","Mada.581","Mada.2226","Mada.118","Mada.1515","Mada.3537","Mada.3707","Mada.4306","Mada.41","Mada.3112","Mada.2842","Mada.2866","Mada.1007","Mada.384","Mada.3162","Mada.2548","Mada.100","Mada.3485","Mada.3666","Mada.1497","Mada.3726","Mada.4301","Mada.4313","Mada.407","Mada.2362","Mada.2339","Mada.3684","Mada.4146","Mada.1865","Mada.2619","Mada.1817","Mada.2608","Mada.2394","Mada.3413","Mada.184","Mada.3674","Mada.2604","Mada.3851","Mada.3796","Mada.1327","Mada.423","Mada.392","Mada.1365","Mada.2605","Mada.2046","Mada.1986","Mada.1251","Mada.1254","Mada.1670","Mada.2437","Mada.3496","Mada.2814","Mada.1905","Mada.614","Mada.2325","Mada.2641","Mada.2401","Mada.3695","Mada.4007","Mada.1852","Mada.3209","Mada.2850","Mada.1132","Mada.2301","Mada.2367","Mada.2756","Mada.715","Mada.1965","Mada.3383","Mada.2387","Mada.3034","Mada.1716","Mada.3326","Mada.2591","Mada.2637")

tree_order_all = c(tree_order, tree_order2)

# reversing order because the text string was top to bottom and tree/ bar plot is made bottom to top.
tree_order_all_rev = rev(tree_order_all)

barchart.full.mada517MB <- mada517.longMB %>%
  ggplot(aes(x = percentage, 
             y = factor(Isolate.IDb, levels = tree_order_all_rev), fill = type)) + ## ** this line is key
  #ggplot(aes(x = percentage, y = as.factor(OrderNum), fill = type)) +
  geom_bar(stat = "identity", position = "stack", width = 0.5) +
  ggthemes::scale_fill_colorblind()+
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_text(size = 10),
    legend.key.size = unit(2, "lines"),
    legend.title = element_text(size = 25),
    legend.text = element_text(size = 25)
  )


barchart.full.mada517MB
#ggsave("./517MB_barchart_factorlevels.pdf", barchart.full.mada517MB, height = 150, width = 30, dpi = 600, limitsize = FALSE)

MadaDB_VsearchClusters_wdata_nest = MadaDB_VsearchClusters_wdata_nest %>%
  mutate(TotalIso = map(data,~ nrow(.x))) %>% # filter clusters with 1 isolate
  filter(TotalIso > 2)

MadaDB_VsearchClusters_wdata_nest = MadaDB_VsearchClusters_wdata_nest %>%
  mutate(FunctionCount = map(data,summary_fxn)) %>% # this creates the function categories
  unnest(FunctionCount) # this allows the function categories to be viewed in the nested frame and "operated" on
 
## This is removing clusters that doesn't have mixed function, meaning inhibitory and not inhibitory
MadaDB_VsearchClusters_wdata_nest = MadaDB_VsearchClusters_wdata_nest %>%
   drop_na(Inhibitory,NonInhibitory)

tbj.77 = MadaDB_VsearchClusters_wdata_nest

tbj.77labels <- tbj.77 %>% 
  tidyr::unnest(data) %>% 
  dplyr::filter(Id.Match != "*") %>% 
  dplyr::group_by(Id.Match) %>% 
  dplyr::slice_head(n = 1) %>% 
  dplyr::ungroup()


ft2.77.data <- full.tree2$data

label.data.77 <- ft2.77.data %>% 
  filter(isTip & label %in% tbj.77labels$Id.Match)



#plotting with the tip labels created before
fulltree2.wtiplabs <- full.tree2.1 +
  ggnewscale::new_scale_color() +   # NEW color scale for tip labels
  geom_tiplab(
    data = label.data.77,
    aes(label = "*", color = `anti-Bd Class (80%)`),
    offset = 0.5,
    size = 6
  ) +
  scale_color_manual(
    values = c(
      "Inhibitory" = "green",
      "NonInhibitory" = "red"
    ),
    name = "Inhibition Status"
  )

fulltree2.wtiplabs


ggsave("recolored_517_wtipstars.pdf", fulltree2.wtiplabs, height = 90, width = 30, dpi = 600, limitsize = FALSE)
```
