---
title: "CultureData_phylogenetics"
author: "Talon Jost"
date: "2025-06-17"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## packages

```{r cars}
rm(list = ls())
library(tidyverse)
```

## Madagascar Culture Metadata input

```{r}

## merging Madagascar database metadata (culture info, etc) with Vsearch 100% clusterIDs

## sorted by seq plate, toxonomy string then isolate ID in excel before creation

MadaDBdata = read_tsv("./MadaCultureDB_Metadata.txt")

## This deals with the duplicate entries for the isolates; its not perfect, it just selects the first entry from the sorted file 

MadaDBdata_filt = MadaDBdata %>%
  group_by(Isolate.ID) %>% slice_head() %>% # filters to 1 sequence per group
  ungroup() %>%
  select(Isolate.ID,Frog.ID,`proportion growth 1`,`anti-Bd Class (80%)`,Taxnonomy,DirTax_Phylum,DirTax_Class,DirTax_Order,DirTax_Family,DirTax_Genus,DirTax_Species) # selecting only a subset of columns for now for easier viewing

```

## Vsearch data

```{r}
#(vsearch) mcbletz@Mollys-MacBook-Pro-2 ~ % vsearch --cluster_fast /Users/mcbletz/Library/CloudStorage/OneDrive-ThePennsylvaniaStateUniversity/Bletz_Lab/Computational_Resources/Datasets/CultureData/MadaCultureDB_seqs_dupsfilt.fasta --uc /Users/mcbletz/Library/CloudStorage/OneDrive-ThePennsylvaniaStateUniversity/Bletz_Lab/Computational_Resources/Datasets/CultureData/MadaCultureDB_seqs_clustered_per100_050725.txt --id 1 --strand both

##*** MODIFIED HEADERS AND COLUMNS MANUALLY IN EXCEL

## Vsearch done with fasta from MadaDB file filtered for duplicates; this has more isolates than AmphiBac; This is the vsearch output (code above); Dup filtering completed with MadaDB_VsearchClusterMergeCode.R file.

MadaDB_VsearchClusters2 = read_tsv("./MadaCultureDB_seqs_clustered_per100_050725.txt") %>%
  select(Record,Cluster,Isolate.ID,Id.Match) # selecting relevant columns


```

## Merging vsearch clusters with culture metadata

Isolate ID is the common identifier and is used to merge the two files

```{r}

## merging clustering with Mada DB metadata that has had the dubs filtered
## NOTE Vsearch was not done with the duplicate filtered fasta...I think that's OK for now.but it does mean that there could be isolates with same ID that are "different" clusters because there are some that didn't ID to the same taxonomy.
MadaDB_VsearchClusters_wdata = left_join(MadaDB_VsearchClusters2,
                                         MadaDBdata_filt,
                                         by = "Isolate.ID") 
#write_tsv(MadaDB_VsearchClusters_wdata,"./MadaCultureDB_Metadata_wVsearchClusters.txt")
```

## Nesting

Nesting groups the Vsearch clusters for further summarizing

IMPORTANT - THERE ARE actually 1387 unique clusters when including singletons!!
There are 817 when we apply the "extreme" function filter

```{r}

MadaDB_VsearchClusters_wdata_nest = MadaDB_VsearchClusters_wdata %>%
  filter(!Record=="C") %>% # removing these records that are not important for the clustering its a duplicate of the S ids; only "S" and "H" retained
  filter(`anti-Bd Class (80%)` == "Inhibitory" | `anti-Bd Class (80%)` == "NonInhibitory") %>%
  filter(`proportion growth 1` > 0.75 | `proportion growth 1` < 0.25) %>% #added this to filter to extreme function
  group_by(Cluster) %>%
  nest()



#need this to have the full list of ids to make the tree
MadaDB_VsearchClusters_wdata_nest.singleton = MadaDB_VsearchClusters_wdata %>%
  filter(!Record=="C") %>% # removing these records that are not important for the clustering its a duplicate of the S ids; only "S" and "H" retained
  filter(`anti-Bd Class (80%)` == "Inhibitory" | `anti-Bd Class (80%)` == "NonInhibitory") %>%
  #filter(`proportion growth 1` > 0.75 | `proportion growth 1` < 0.25) %>% #added this to filter to extreme function
  group_by(Cluster) %>%
  nest()


##matching this to where I can actually get the isolate IDs to match the cluster.
singelton.ids <- MadaDB_VsearchClusters_wdata_nest.singleton$Cluster
singleton.df <- MadaDB_VsearchClusters2 %>% 
  dplyr::filter(Cluster %in% singelton.ids) %>% 
  dplyr::group_by(Cluster) %>% 
  dplyr::slice_head(n = 1)


singleton.df2 <- MadaDB_VsearchClusters_wdata_nest.singleton %>% 
  dplyr::mutate(IsSingleton = purrr::map_int(data, nrow) == 1)

singleton.df2 <- singleton.df2 %>% 
  dplyr::left_join(MadaDB_VsearchClusters2, by = "Cluster") %>% 
  dplyr::filter(!Record=="C") %>% 
  dplyr::group_by(Cluster) %>% 
  dplyr::slice_head(n = 1)

singleton.df3 <- singleton.df2 %>% 
  dplyr::left_join(MadaDBdata_filt, by = "Isolate.ID") %>% 
  dplyr::rename(Inhibition = `anti-Bd Class (80%)`)

singleton.df3b <- singleton.df3 %>% 
  dplyr::rename(tip.label = Isolate.ID) %>% 
  dplyr::select(Cluster, IsSingleton, tip.label) 



MadaDBdata_filt <- MadaDBdata_filt %>% 
  dplyr::left_join(singleton.df3, by = "Isolate.ID")
```

## Summarizing

This creates an initial summary of the functional categories for each cluster


```{r}

summary_fxn = function(x){
  x %>% ## called x because thats what the position in the map function is called
  group_by(`anti-Bd Class (80%)`) %>%
  summarize(Count = n()) %>%
    pivot_wider(.,names_from = `anti-Bd Class (80%)`,values_from = Count)

} # this is a function that does the summarizing by function category and then organizes it for better viewing.

#MadaDB_VsearchClusters_wdata_nest = MadaDB_VsearchClusters_wdata_nest %>%
 # mutate(TotalIso = map(data,~ nrow(.x)))

MadaDB_VsearchClusters_wdata_nest = MadaDB_VsearchClusters_wdata_nest %>%
  mutate(TotalIso = map(data,~ nrow(.x)))%>%
  mutate(FunctionCount = map(data,summary_fxn)) %>% # this creates the function categories
  unnest(FunctionCount) # this allows the function categories to be viewed in the nested frame and "operated" on


MadaDB_VsearchClusters_wdata_nest_nosingles = MadaDB_VsearchClusters_wdata_nest %>%
  filter(TotalIso > 1) %>% # filter clusters with 1 isolate
  drop_na(Inhibitory,NonInhibitory) %>%
  mutate(InhibitoryRatio = paste(Inhibitory,".",NonInhibitory))
  
  
  ## List of IDs for tree for all cluster dataset
MadaDB_VsearchClusters_wdata_nestnosingles_IDs = MadaDB_VsearchClusters_wdata_nest517 %>%
  unnest(data) %>%
  select(Id.Match) %>%
  filter(!Id.Match == "*") %>% ## This is needed to remove the creation of an "asterisk" group
  unique()

### STOPPED HERE DURING MEETING on 17 June 2025 ############## 



#MadaDB_VsearchClusters_wdata_nest2 = MadaDB_VsearchClusters_wdata_nest %>%
 # mutate(TotalIso = map(data,~ nrow(.x))) %>% # filter clusters with 1 isolate
  #filter(TotalIso > 2) %>% 
  #unnest(data) %>%
  #group_by(DirTax_Phylum) %>%
  #summarize(Count = n_distinct(Cluster))

MadaDB_VsearchClusters_wdata_nest = MadaDB_VsearchClusters_wdata_nest %>%
  mutate(FunctionCount = map(data,summary_fxn)) %>% # this creates the function categories
  unnest(FunctionCount) # this allows the function categories to be viewed in the nested frame and "operated" on
 
## This is removing clusters that doesn't have mixed function, meaning inhibitory and not inhibitory
 MadaDB_VsearchClusters_wdata_nest = MadaDB_VsearchClusters_wdata_nest %>%
   drop_na(Inhibitory,NonInhibitory)
 

 

```



####plot the tree
```{r plot, echo = TRUE}
library(ape)
tree.original <- ape::read.tree('trees/tree/nwk/madaDB.rooted.tree.nwk')
#plot(tree.original)

names.to.keep <- singleton.df3b$tip.label
keep <- intersect(names.to.keep, tree.original$tip.label)

tree.pruned <- keep.tip(tree.original, keep)
tree.pruned$tip.label
plot(tree.pruned)

library(ggtree)
library(ggtreeExtra)
library(treeio)


tree.original.ggtree.517 <- ggtree(tree.pruned, branch.length = "none", size = 0.5) %<+% MadaDBdata_filt +
  geom_tree(aes(color = DirTax_Class.x)) +
    scale_color_manual(name = "Class", values = c(
  "Actinobacteria" = "#4E79A7",
  "Deinococci" = "#F28E2B",
  "Alphaproteobacteria" = "#E15759",
  "Gammaproteobacteria" = "#76B7B2",
  "Betaproteobacteria" = "#59A14F",
  "Flavobacteriia" = "#EDC948",
  "Sphingobacteriia" = "#B07AA1",
  "Bacilli" = "#FF9DA7",
  "[Saprospirae]" = "#ff833a",
  "Cytophagia" = "#a3b"
), na.value = "grey") +
  ggnewscale::new_scale_color() +
  geom_tippoint(aes(color = IsSingleton), shape = 16, size = 8) +
  scale_color_manual(values = c(
    "TRUE" = "green",
    "FALSE" = "black",
    "NA" = "grey"
  ), name = "Inhibition Status") +
  geom_tiplab() +
  theme_tree()
tree.original.ggtree.517

#ggsave("singleton_517_tree2.pdf", tree.original.ggtree.517, height = 400, width = 50, dpi = 800, limitsize = FALSE)
```



#barchart
```{r}
#id list copied directly from the plot above
ids <- readLines("list_names.txt")
ids <- unique(ids)
#formatted <- paste0('"', ids, '",')
#writeLines(formatted, "list_names_quoted.txt")


# 
# ordering <- match(ids, names.to.keep)
# sorted_lines <- ids[ordering]
# 
# writeLines(sorted_lines, "file_sorted.txt")

tip.labels.order <- tree.pruned$tip.label
identical(tip.labels.order, ids)

bar.data <- MadaDB_VsearchClusters_wdata_nest.singleton %>% 
  mutate(
    TotalIso = map_int(data, nrow),
    Inhibitory = map_int(data, ~ sum(.x$`anti-Bd Class (80%)` == "Inhibitory", na.rm = TRUE)),
    NonInhibitory = map_int(data, ~ sum(.x$`anti-Bd Class (80%)` == "NonInhibitory", na.rm = TRUE)),
    Null = map_int(data, ~ sum(is.na(.x$`anti-Bd Class (80%)`))),
    Contam = map_int(data, ~ sum(.x$`anti-Bd Class (80%)` == "Contam", na.rm = TRUE))
  )

bar.data2 <- bar.data %>% 
  dplyr::left_join(MadaDB_VsearchClusters_wdata, by = "Cluster") %>% 
  dplyr::select(Cluster, data, Isolate.ID, 
                TotalIso, NonInhibitory, Inhibitory, 
                Null, Contam) %>% 
  dplyr::group_by(Cluster) %>% 
  dplyr::slice_head(n = 1)


bar.data3 <- bar.data2 %>% 
  dplyr::mutate(
    prop.inhib = Inhibitory / TotalIso * 100,
    prop.noninhib = NonInhibitory / TotalIso * 100,
    prop.na = Null / TotalIso * 100,
    prop.contam = Contam / TotalIso * 100
  )

bar.long <- bar.data3 %>%
  select(Isolate.ID, Inhibitory, NonInhibitory, prop.inhib, prop.noninhib, TotalIso) %>%
  pivot_longer(cols = c(prop.inhib, prop.noninhib),
               names_to = "type",
               values_to = "percentage")

# Recode type labels
bar.long$type <- dplyr::recode(bar.long$type,
                               prop.inhib = "Inhibitory",
                               prop.noninhib = "Non-Inhibitory")


levels(bar.long$Isolate.ID)

# Build bar.label.df if needed
bar.label.df <- bar.data3 %>%
  mutate(label = paste0("n=", TotalIso),
         xpos = prop.inhib + prop.noninhib) %>%
  select(Isolate.ID, xpos, label)

barchart.108 <- bar.long %>%
  ggplot(aes(x = percentage, 
             y = factor(Isolate.ID, levels = ids), 
             fill = type)) + ## ** this line is key
  geom_bar(stat = "identity", 
           position = "stack", 
           width = 0.5) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_text(size = 10),
    legend.key.size = unit(2, "lines"),
    legend.title = element_text(size = 25),
    legend.text = element_text(size = 25)
  )

barchart.108

#ggsave("barchart108.pdf", barchart.108, height = 400, width = 50, limitsize = FALSE)
```

no singles
```{r}
MadaDB_VsearchClusters_wdata_nestnosingles_ids2 <- MadaDB_VsearchClusters_wdata_nest_nosingles %>% dplyr::select(Cluster)

nest108 <- MadaDB_VsearchClusters_wdata_nest_nosingles

nest108.1 <- nest108 %>% 
  tidyr::unnest(data) %>% 
  dplyr::group_by(Id.Match) %>% 
  dplyr::slice_head(n = 1) %>% 
  dplyr::filter(Id.Match != "*") #filter this as not having it will result in one more observation that already exists within the dataframe

nest108.1.Ids <- nest108.1$Id.Match

madafilt108 <- MadaDBdata_filt %>% 
  dplyr::mutate(IsCluster = Isolate.ID %in% nest108.1.Ids)

join108 <- nest108.1 %>% 
  dplyr::select(Isolate.ID, InhibitoryRatio)

madafilt108.1 <- madafilt108 %>% 
  dplyr::left_join(join108, by = "Isolate.ID")

highlight_nodes <- data.frame(
  node = c(2405, 2392, 2387, 2349,
           2324, 2193, 2163, 2136,
           2098, 1891, 1903, 1810,
           1794, 1824, 1854, 1969,
           1947, 2020, 2012, 2059,
           1753, 1748, 1648, 1579,
           1685, 1670, 1465, 1457,
           1487, 1511, 1533, 1547,
           1540, 1382, 1405, 1341,
           1307, 1265, 1232, 1208),
  group = paste0("Cluster_", make.unique(rep(LETTERS, length.out = 40)))
)

node_plot_data <- dplyr::inner_join(tree.data, node_colors, by = "node")

tree.original.ggtree.108 <- ggtree(tree.pruned, branch.length = "none", size = 0.5) %<+% madafilt108.1 +
  geom_tree(aes(color = DirTax_Class.x)) +
    scale_color_manual(name = "Class", values = c(
  "Actinobacteria" = "#4E79A7",
  "Deinococci" = "#F28E2B",
  "Alphaproteobacteria" = "#E15759",
  "Gammaproteobacteria" = "#76B7B2",
  "Betaproteobacteria" = "#59A14F",
  "Flavobacteriia" = "#EDC948",
  "Sphingobacteriia" = "#B07AA1",
  "Bacilli" = "#FF9DA7",
  "[Saprospirae]" = "#ff833a",
  "Cytophagia" = "#a3b"
), na.value = "grey") +
  ggnewscale::new_scale_color() +
  geom_tippoint(aes(color = IsCluster), shape = 16, size = 8) +
  scale_color_manual(values = c(
    "TRUE" = "green",
    "FALSE" = "black",
    "NA" = "grey"
  ), name = "Is 108") +
  geom_tiplab() +
  ggnewscale::new_scale_color() +
  geom_tiplab(aes(color = IsSingleton)) +
  scale_color_manual(values = c(
    "TRUE" = "cyan3",
    "FALSE" = "black",
    "NA" = "grey"
  ), name = "Is Singleton") +
  theme_tree() #+
  #geom_text2(aes(label = node), hjust = -0.3, size = 3)
tree.original.ggtree.108

ggsave("singleton_108_tree.pdf", tree.original.ggtree.108, height = 400, width = 50, dpi = 800, limitsize = FALSE)

tree.data <- ggtree::fortify(tree.pruned)

tree.original.ggtree.108 <- tree.original.ggtree.108 +
  ggtree::geom_hilight(
    data = highlight_nodes,
    mapping = aes(node = node, fill = group),
    type = "roundrect",
    alpha = 0.3,
    extend = 0.5
  ) +
  scale_fill_manual(values = setNames(rep("red", length(unique(highlight_nodes$group))), unique(highlight_nodes$group)),
                    name = "Highlighted Nodes")

ggsave("singleton_108_tree.pdf", tree.original.ggtree.108, height = 400, width = 50, dpi = 800, limitsize = FALSE) 
```

```{r}
tree.original.ggtree.108.rev <- ggtree(tree.pruned, branch.length = "none", size = 0.5) %<+% madafilt108 +
  geom_tree(aes(color = DirTax_Class.x)) +
    scale_color_manual(name = "Class", values = c(
  "Actinobacteria" = "#4E79A7",
  "Deinococci" = "#F28E2B",
  "Alphaproteobacteria" = "#E15759",
  "Gammaproteobacteria" = "#76B7B2",
  "Betaproteobacteria" = "#59A14F",
  "Flavobacteriia" = "#EDC948",
  "Sphingobacteriia" = "#B07AA1",
  "Bacilli" = "#FF9DA7",
  "[Saprospirae]" = "#ff833a",
  "Cytophagia" = "#a3b"
), na.value = "grey") +
  ggnewscale::new_scale_color() +
  geom_tippoint(aes(color = IsCluster), shape = 16, size = 8) +
  scale_color_manual(values = c(
    "TRUE" = "blue",
    "FALSE" = "black",
    "NA" = "grey"
  ), name = "Is 108") +
  theme_tree()
tree.original.ggtree.108.rev

ggsave("singleton_108_tree.rev.pdf", tree.original.ggtree.108.rev, height = 400, width = 50, dpi = 800, limitsize = FALSE)
```

