---
title: "mada85_flux_analysis"
author: "Talon B Jost"
date: "2025-09-03"
output: pdf_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
##analysing the data some more and making figures
library(sybil)
library(sybilSBML)
library(glpkAPI)
library(ggplot2)
```


#systems biology markup language interpreter
#this is taking the metabolic reactions created by gapseq and pulling them into the environment
#this reads the SBML files to use in constraint based modeling
#SBML describes:
# - Species (metabolites, proteins, RNA)
# - Reactions (biochemical reactions or transport processes)
# - Compartments (cytoplasm, mitochondria, etc.)
# - Kinetic laws or constraints (reaction rates, bounds)
# - Gene-reaction associations
```{r}
#what this is doing is constraint based metabolic modeling or flux balance analysis
#flux balance analysis is the 
model85 <- sybilSBML::readSBMLmod("gapseq.mada85/Mada85.xml")

#check the model summary
model85

#optimize the model for the entire components
result <- optimizeProb(model85)

sybil::react_name(model85)


#check flux balance analysis
# Objective value (e.g., growth rate)
# the objective value is the value of the objective function (usually growth rate) at the optimal solution given all constraints.
print(paste("Objective value:", result@lp_obj))


# Convert
flux_vector <- fluxes(result)

# Convert sparse matrix to numeric vector
# sparse vector just saves on space
flux_vector_num <- as.numeric(flux_vector)

# Now create data frame
flux_df <- data.frame(
  reaction = react_id(model),
  flux = flux_vector_num
)

head(flux_df)

```


making a figure from it
```{r}
ggplot(flux_df, aes(x = flux)) +
  geom_histogram(bins = 50, fill = "steelblue", color = "black") +
  theme_minimal() +
  labs(title = "Flux Distribution",
       x = "Flux value",
       y = "Number of reactions")

```

Reading and interpreting this figure
1. Network sparsity and activity
- In a metabolic network, not all reactions are active at once under a given condition (e.g., minimal medium).
- Reactions with zero flux are inactive — the cell doesn’t need them to achieve the objective (biomass growth).
- Only a subset of reactions carry flux — this is typical in constraint-based models.

2. High flux reactions
- Reactions at the extreme ends (high positive or negative flux) are key active pathways:
- Positive flux → forward reaction carrying material toward biomass or product formation.
- Negative flux → reverse reaction (if reversible).
- These are often bottleneck or essential reactions for growth.

3. Biological insight
- The histogram tells you the network is highly selective:
- The organism only uses a few reactions efficiently to achieve growth.
- Many reactions are theoretically possible but not used under this condition.
- This is a common feature of FBA because it finds one optimal flux distribution that maximizes biomass.


Single gene knockouts doesn't really work.
```{r}
#single gene knockouts
library(sybil)
library(sybilSBML)

# -----------------------------
# Load the model
# -----------------------------
model <- readSBMLmod("gapseq.mada85/Mada85.xml")

# -----------------------------
# Identify biomass reaction
# -----------------------------
biomass_candidates <- grep("biomass|BIOMASS", react_name(model), ignore.case = TRUE)
if(length(biomass_candidates) == 0) stop("No biomass reaction found. Please check your model.")
biomass_rxn <- react_id(model)[biomass_candidates[1]]
cat("Using biomass reaction:", biomass_rxn, "\n")

# Set biomass as the objective
obj_coef(model) <- rep(0, react_num(model))
obj_coef(model)[which(react_id(model) == biomass_rxn)] <- 1

# -----------------------------
# Open exchange reactions (uptake)
# -----------------------------
Smat <- S(model)
ex_idx <- which(colSums(abs(Smat) > 0) == 1)  # single metabolite reactions
lowbnd(model)[ex_idx] <- -10  # allow uptake

# -----------------------------
# Wild-type growth
# -----------------------------
wt <- optimizeProb(model)
wildtype <- wt@lp_obj
cat("Wild-type growth:", wildtype, "\n")
if(wildtype <= 0) warning("Wild-type growth is zero. Check biomass reaction and medium.")

# -----------------------------
# Helper functions
# -----------------------------
getReactionsForGene <- function(model, gene) {
  gpr_rules <- gprRules(model)
  gene_idx <- match(gene, allGenes(model))
  if(is.na(gene_idx)) return(integer(0))
  pattern <- paste0("x\\[", gene_idx, "\\]")
  which(grepl(pattern, gpr_rules))
}

knockoutGene <- function(model, gene) {
  rxns <- getReactionsForGene(model, gene)
  ko_model <- model
  if(length(rxns) > 0) {
    lowbnd(ko_model)[rxns] <- 0
    uppbnd(ko_model)[rxns] <- 0
  }
  return(ko_model)
}

# -----------------------------
# Identify genes with at least 1 reaction
# -----------------------------
gene_list <- allGenes(model)
rxn_counts <- sapply(gene_list, function(g) length(getReactionsForGene(model, g)))
functional_genes <- gene_list[rxn_counts > 0]
cat("Functional genes (have reactions):", length(functional_genes), "\n")

# -----------------------------
# Single-gene knockout loop
# -----------------------------
ko_results <- numeric(length(functional_genes))
names(ko_results) <- functional_genes

for(g in functional_genes){
  ko_model <- knockoutGene(model, g)
  ko_sol <- optimizeProb(ko_model)
  ko_results[g] <- ko_sol@lp_obj
}

# -----------------------------
# Summarize results
# -----------------------------
ko_df <- data.frame(
  gene = functional_genes,
  growth_after_KO = ko_results,
  rel_growth = ko_results / wildtype
)

# Sort by growth after KO
ko_df <- ko_df[order(ko_df$growth_after_KO), ]
head(ko_df, 10)


```


#find reaction level differences using gapseq files
```{r}
# Load necessary libraries
library(dplyr)
library(readr)
library(tidyr)

setwd("~/Documents/Penn_State/Genomes/SynCom")

# ------------------------------
# Step 1: Read gapseq annotation files
# ------------------------------
mada85 <- read.delim("gapseq.mada85/Pathways.edited.tsv")
mada106 <- read.delim("gapseq106/Pathways106.edited.tsv")


# ------------------------------
# Step 2: merge files
# ------------------------------

p.comp <- merge(mada85, mada106, by = "ID", suffixes =c(".g1", ".g2"))

#subset to unique for both dataframes
unique.85g1 <- subset(p.comp, Completeness.g1 > 0 & Completeness.g2 == 0)
unique.106g2 <- subset(p.comp, Completeness.g1 == 0 & Completeness.g2 > 0)
shared <- subset(p.comp, Completeness.g1 > 0 & Completeness.g2 > 0)

# ------------------------------
# Step 3: Identify unique and shared genes
# ------------------------------
print(unique.85g1$Name.g1)
print(unique.106g2$Name.g2)



#venn diagram
library(VennDiagram)
library(grid)

set1 <- p.comp$Name.g1[p.comp$Completeness.g1 > 0]
set2 <- p.comp$Name.g2[p.comp$Completeness.g2 > 0]

only85 <- setdiff(set1, set2)
only106 <- setdiff(set2, set1)
both <- intersect(set1, set2)

venn.plot <- venn.diagram(
  x = list("Mada.85: NonInhibitory" = set1, "Mada.106: Inhibitory" = set2),
  filename = NULL,
  fill = c("skyblue", "lightgreen"),
  alpha = 0.5,
  scaled = FALSE,
  cex = 0,        # text size for numbers (not used here, can keep small)
  cat.cex = 1.5,  # title size
  cat.pos = c(180, 180),  # angle of titles: first circle left, second circle right
  cat.dist = c(0.05, 0.05), # distance of titles from circles
  margin = 0.1
)

grid.newpage()
grid.draw(venn.plot)

grid.text(paste(only85, collapse = "\n"), x = 0.25, y = 0.55, gp = gpar(cex = 0.9))
grid.text(paste(only106, collapse = "\n"), x = 0.75, y = 0.55, gp = gpar(cex = 0.9))


install.packages("UpSetR")
library(UpSetR)

input <- list(
  Genome85 = set1,
  Genome106 = set2
)

upset(fromList(input), nsets = 2, order.by = "freq")


# Shared genes
shared_genes <- mada85 %>%
  filter(Gene %in% mada106$Gene)

# ------------------------------
# Step 4: Optional - Combine for comparison table
# ------------------------------
comparison <- full_join(mada85 %>% select(Gene, Product, COG) %>% rename(Mada85_Product = Product, Mada85_COG = COG),
                        mada106 %>% select(Gene, Product, COG) %>% rename(Mada106_Product = Product, Mada106_COG = COG),
                        by = "Gene")

# ------------------------------
# Step 5: Save results
# ------------------------------
write_csv(mada85_unique, "mada85_unique_genes.csv")
write_csv(mada106_unique, "mada106_unique_genes.csv")
write_csv(shared_genes, "shared_genes.csv")
write_csv(comparison, "gene_comparison_table.csv")
```

