---
title: "mada106_gapseq"
author: "Talon B Jost"
date: "2025-09-08"
output: pdf_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
##analysing the data some more and making figures
library(sybil)
library(sybilSBML)
library(glpkAPI)
library(ggplot2)
```


#systems biology markup language interpreter
#this is taking the metabolic reactions created by gapseq and pulling them into the environment
#this reads the SBML files to use in constraint based modeling
#SBML describes:
# - Species (metabolites, proteins, RNA)
# - Reactions (biochemical reactions or transport processes)
# - Compartments (cytoplasm, mitochondria, etc.)
# - Kinetic laws or constraints (reaction rates, bounds)
# - Gene-reaction associations
```{r}
#what this is doing is constraint based metabolic modeling or flux balance analysis
#flux balance analysis is the 
model106 <- sybilSBML::readSBMLmod("gapseq106/fluxanalysis/Mada106.xml")

#check the model summary
model106

#optimize the model for the entire components
result106 <- optimizeProb(model106)

sybil::react_name(model106)


#check flux balance analysis
# Objective value (e.g., growth rate)
# the objective value is the value of the objective function (usually growth rate) at the optimal solution given all constraints.
print(paste("Objective value:", result106@lp_obj))
#these are different between 106 and 85

# Convert
flux_vector106 <- fluxes(result106)

# Convert sparse matrix to numeric vector
# sparse vector just saves on space
flux_vector_num106 <- as.numeric(flux_vector106)

# Now create data frame
flux_df106 <- data.frame(
  reaction = react_id(model106),
  flux = flux_vector_num106
)

head(flux_df106)

```


making a figure from it
```{r}
ggplot(flux_df106, aes(x = flux)) +
  geom_histogram(bins = 50, fill = "steelblue", color = "black") +
  theme_minimal() +
  labs(title = "Flux Distribution 106",
       x = "Flux value",
       y = "Number of reactions")

```

Reading and interpreting this figure
1. Network sparsity and activity
- In a metabolic network, not all reactions are active at once under a given condition (e.g., minimal medium).
- Reactions with zero flux are inactive — the cell doesn’t need them to achieve the objective (biomass growth).
- Only a subset of reactions carry flux — this is typical in constraint-based models.

2. High flux reactions
- Reactions at the extreme ends (high positive or negative flux) are key active pathways:
- Positive flux → forward reaction carrying material toward biomass or product formation.
- Negative flux → reverse reaction (if reversible).
- These are often bottleneck or essential reactions for growth.

3. Biological insight
- The histogram tells you the network is highly selective:
- The organism only uses a few reactions efficiently to achieve growth.
- Many reactions are theoretically possible but not used under this condition.
- This is a common feature of FBA because it finds one optimal flux distribution that maximizes biomass.


Single gene knockouts doesn't really work.
